Marlowe: financial contracts on blockchain

### 2018-05-17 tex/paper.tex
moved related work to the end; edits/changes to the intro

In doing this, we are required to refine the model of contracts in a number of ways in order to fit with a radically 
different context. Let us consider the following example of an ``escrow'' contract so that we can explain the motivation 
more concretely. The aim of this contract, written in functional pseudocode in the style of~\cite{PeytonJones:2000} 
involves three participants \haskellinline{1}, \haskellinline{2} and \haskellinline{3}. Participant \haskellinline{1} is 
to pay an amount of money to participant \haskellinline{2} on receipt of goods from her. Participant \haskellinline{1} 
pays the money into escrow controlled by participant \haskellinline{3}. There are two options for the money: if two 
participants agree to \haskellinline{pay} it to \haskellinline{2}, that goes ahead; if, on the other hand, two of the 
participants opt to \haskellinline{refund} the money to \haskellinline{1}, that is done instead. The 
\haskellinline{When} construct fires when its first argument -- a condition -- becomes true; the second argument, a 
\haskellinline{Contract} itself, is then performed, and this makes the payment according to the participants' 
\haskellinline{Choice}.

@@

In the traditional model, enforcement of the contract is the responsibility of the legal system. If participant 
\haskellinline{1} does not pay the money into escrow, or participant \haskellinline{3} chooses to keep it for herself, 
then we can assume that they can be sued for the money and probably damages on top, thus providing both legal and 
financial incentives for compliance. On the other hand, in the decentralised blockchain model, where there is no central 
authority, the contract needs to be enforced by design.

This means that we require participants to commit money to cover all possible expenditure \emph{in advance of the 
contract executing}. In order to make sure that participants continue to engage with a contract, we ensure urgency by 
imposing timeouts: money is committed for a finite period, for example, and when waiting for a participant to make a 
commitment we also impose a timeout to ensure that the contract does not become stuck even if one or more of the 
participants stop interacting with it. We revisit this example in Section~\ref{section:example-escrow} where we show how 
commitments, timeouts and other features of Marlowe are used to describe this as a self-enforcing contract ``on the 
blockchain''.

Moreover, there is not a single blockchain model, and we give a brief overview of the different mechanisms for 
scripting on blockchain (although this is not the primary purpose of this article). 
A particularly important technical distinction here is that between an account-style model as used by 
Ethereum~\cite{EthereumRationale}, and the unspent transaction output (UTxO) model of bitcoin~\cite{sok}. 
There is also the question of how precisely the execution of the contract interacts with the blockchain, and how 
information is to be recorded on the chain. Finally, should the execution ``push'' for changes  to happen, or do 
external actors have to ``pull'' to effect changes enabled by execution of the contract?

The paper describes version 1.3 of Marlowe, and it should be seen as a work in progress. We have designed it to provide 
financial contracts in Cardano, but also in order to start to answer the twin questions of what smart contracts will 
look like in Cardano, and how those contracts will be implemented in the overall system.

The paper is structured like this: we begin in Section \ref{section:design} with a discussion of the design rationale 
for Marlowe, including an overview of the various approaches to scripting in blockchain as well as the different designs 
of the blockchains themselves. We conclude Section \ref{section:design} with a set of decisions about the blockchain 
environment for Marlowe. The central part of the paper is Section \ref{section:model} where we introduce the Marlowe 
model, including the assumptions made in designing it, the types of the principal functions and a description of the DSL 
as an algebraic type, constructor by constructor. 
%Section \ref{section:example-saving} presents an example of a larger Marlowe contract, for incentivised saving. 
Section \ref{section:example-escrow} revisits the escrow example, and shows 
how it is described using a combination of Marlowe and Haskell constructs: that is, we use Marlowe as an embedded DSL. 
Section \ref{section:tool} introduces Meadow, our tool for visualising and interacting with Marlowe contracts.
Section \ref{section:reflection} reflects on the system design, and observes some alternative designs and Section 
\ref{section:next-steps} enumerates the next steps for the project.

@@

\subsection{Smart contracts on blockchains}

As more than one wag has observed, many of the programs that run on blockchains are neither smart nor indeed form 
contracts in any real sense. Nonetheless, this terminology has stuck, and has become shorthand for ``programmed 
interactions on the blockchain''. In an earlier paper, we presented a survey of the approaches 
available~\cite{cryptoeprint:2016:1156}; we give an overview here.

Blockchains are executed in a replicated form by parties who cannot be guaranteed not to be hostile, either by directly 
trying to change the contents of the chain, or through trying to affect other properties of the chain by indirect means 
(such as swamping honest parties with work). Programming on the blockchain therefore needs to be constrained in some 
ways, since it has to be amenable to replication or verification within a reasonable time if the security and integrity 
of the chain are to be preserved. We look at representatives of the main approaches now.

\paragraph{Related work}

@@

A running contract will interact with its environment in two ways, as shown in Figure \ref{fig:context}.

@@

\begin{figure}
\centering{}\includegraphics[scale=0.5]{pix/detail2}\caption{\label{fig:detail-of-interface}Detail of the interface for 
contract
execution simulation.}
\end{figure}

Additionally, to facilitate the introduction of inputs, Meadow provides
two different interfaces:
\begin{itemize}
\item The manual interface (see Figure~\ref{fig:detail-of-manual-interface}),
provides a template for each of the four possible types of input:
commits, redeems, payment claims, and choices.
\item The smart interface (see Figure~\ref{fig:detail-of-smart-interface}),
calculates the possible operations that would make sense given the
current inputs, state of the contract, block number, and remaining
contract; and it provides them in a table with most of the parameters
already filled in.
\end{itemize}
\begin{figure}
\begin{centering}
\includegraphics[scale=0.5]{pix/detail3}
\par\end{centering}
\caption{\label{fig:detail-of-manual-interface}Detail of the manual interface.}
\end{figure}

\begin{figure}
\centering{}\includegraphics[scale=0.5]{pix/detail4}\caption{\label{fig:detail-of-smart-interface}Detail of the smart 
interface.}
\end{figure}

@@

@@

Thanks to IOHK staff Manuel Chakravarty, Duncan Coutts, Bernardo David, Charles Hoskinson, Aggelos Kiayias, % Darryl McAdams, 
Bruno Woltzenlogel Paleo, Rebecca Valentine and Phil Wadler for their collaboration, and to colleagues from the Universities of 
Kent and Leicester for their comments on previous presentations of this work.

>>>>

In doing this, we are required to refine the model of contracts in a number of ways in order to fit with a radically 
different context. Consider the following example of an ``escrow'' contract so that we can explain the motivation 
more concretely. The aim of this contract, written in functional pseudocode in the style of~\cite{PeytonJones:2000} 
involves three participants \haskellinline{1}, \haskellinline{2} and \haskellinline{3}. Participant \haskellinline{1} is 
to pay an amount of money to participant \haskellinline{2} on receipt of goods from her. Participant \haskellinline{1} 
pays the money into escrow controlled by participant \haskellinline{3}. There are two options for the money: if two 
participants agree to \haskellinline{pay} it to \haskellinline{2}, that goes ahead; if, on the other hand, two of the 
participants opt to \haskellinline{refund} the money to \haskellinline{1}, that is done instead. The 
\haskellinline{When} construct fires when its first argument -- a condition -- becomes true; the second argument, a 
\haskellinline{Contract} itself, is then performed, and this makes the payment according to the participants' 
\haskellinline{Choice}.

@@

In the traditional model, enforcement of the contract is the responsibility of the legal system. If participant 
\haskellinline{1} does not pay the money into escrow, or participant \haskellinline{3} chooses to keep it for herself, 
then they can be sued for the money (and probably damages), thus providing both legal and 
financial incentives for compliance. On the other hand, in the decentralised blockchain model, where there is no central 
authority, the contract needs to be enforced \emph{by design}.

This means that we require participants to \emph{commit} money to cover all possible expenditure \emph{in advance of the 
contract executing}. In order to make sure that participants continue to engage with a contract, we ensure urgency by 
imposing \emph{timeouts}: money is committed for a finite period only. We also impose a timeout when  waiting for a participant to make a 
commitment to ensure that the contract does not become stuck even if one of the 
participants stops interacting with it. We discuss this particular example in more detail in Marlowe in Section~\ref{section:example-escrow} below.

We make the following contributions in this paper.
\begin{itemize}
\item
The design of the Marlowe language as a Haskell data type; Marlowe is made more expressible through its embedding in Haskell.
\item
An executable, small-step semantics of Marlowe in Haskell.
\item
The Meadow tool that allows users to simulate the operation of Marlowe contracts, including ``embedded'' contracts that use features of Haskell. 
\end{itemize}

%Moreover, there is not a single blockchain model, and we give a brief overview of the different mechanisms for 
%scripting on blockchain (although this is not the primary purpose of this article). 
%A particularly important technical distinction here is that between an account-style model as used by 
%Ethereum~\cite{EthereumRationale}, and the unspent transaction output (UTxO) model of bitcoin~\cite{sok}. 
%There is also the question of how precisely the execution of the contract interacts with the blockchain, and how 
%information is to be recorded on the chain. Finally, should the execution ``push'' for changes  to happen, or do 
%external actors have to ``pull'' to effect changes enabled by execution of the contract?
%
%The paper describes version 1.3 of Marlowe, and it should be seen as a work in progress. We have designed it to provide 
%financial contracts in Cardano, but also in order to start to answer the twin questions of what smart contracts will 
%look like in Cardano, and how those contracts will be implemented in the overall system.

The paper is structured like this: REWRITE we begin in Section \ref{section:design} with a discussion of the design rationale 
for Marlowe, including an overview of the various approaches to scripting in blockchain as well as the different designs 
of the blockchains themselves. We conclude Section \ref{section:design} with a set of decisions about the blockchain 
environment for Marlowe. The central part of the paper is Section \ref{section:model} where we introduce the Marlowe 
model, including the assumptions made in designing it, the types of the principal functions and a description of the DSL 
as an algebraic type, constructor by constructor. 
%Section \ref{section:example-saving} presents an example of a larger Marlowe contract, for incentivised saving. 
Section \ref{section:example-escrow} revisits the escrow example, and shows 
how it is described using a combination of Marlowe and Haskell constructs: that is, we use Marlowe as an embedded DSL. 
Section \ref{section:tool} introduces Meadow, our tool for visualising and interacting with Marlowe contracts.
Section \ref{section:reflection} reflects on the system design, and observes some alternative designs and Section 
\ref{section:next-steps} enumerates the next steps for the project.

@@

@@

A running contract interacts with its environment in two ways, as in Figure \ref{fig:context}.

@@

\begin{figure}[t]
\centering{}\includegraphics[scale=0.5]{pix/detail2}\caption{\label{fig:detail-of-interface}Detail of the interface for 
contract
execution simulation.}
\end{figure}

Additionally, to facilitate the introduction of inputs, Meadow provides
two different interfaces:
\begin{itemize}
\item The manual interface (see Figure~\ref{fig:detail-of-manual-interface}),
provides a template for each of the four possible types of input:
commits, redeems, payment claims, and choices.
\item The smart interface (see Figure~\ref{fig:detail-of-smart-interface}),
calculates the possible operations that would make sense given the
current inputs, state of the contract, block number, and remaining
contract; and it provides them in a table with most of the parameters
already filled in.
\end{itemize}
%\begin{figure}
%\begin{centering}
%\includegraphics[scale=0.5]{pix/detail3}
%\par\end{centering}
%\caption{\label{fig:detail-of-manual-interface}Detail of the manual interface.}
%\end{figure}
%
%\begin{figure}
%\centering{}\includegraphics[scale=0.5]{pix/detail4}\caption{\label{fig:detail-of-smart-interface}Detail of the smart 
%interface.}
%\end{figure}

@@

\section{Related work}

Blockchains are executed in a replicated form by parties who cannot be guaranteed not to be hostile, either by directly 
trying to change the contents of the chain, or through trying to affect other properties of the chain by indirect means 
(such as swamping honest parties with work). Programming on the blockchain therefore needs to be constrained in some 
ways, since it has to be amenable to replication or verification within a reasonable time if the security and integrity 
of the chain are to be preserved. We look at representatives of the main approaches now; we discuss some of these in more detail in an earlier paper~\cite{cryptoeprint:2016:1156}.

\paragraph{Findel}

@@

Thanks to Manuel Chakravarty, Duncan Coutts, Bernardo David, Charles Hoskinson, Aggelos Kiayias, % Darryl McAdams, 
Bruno Woltzenlogel Paleo, Rebecca Valentine and Phil Wadler for their collaboration, and to colleagues from the Universities of 
Kent and Leicester for their comments on previous presentations of this work.
